\section{Conclusion}
\label{sec:conclusion}
In this paper, we jointly considered the use of WiFi and white space bands application
for wireless networks deployments. 
Different from prior work, we first proposed a Multiband 
Access Point Estimation (MAPE) framework to estimate the number of access points required in 
a given region for wireless access network and Band-based Path Selection (BPS) algorithm for
backhual network based on in-field measurements for wireless access network. 
We then performed spectrum utilization measurements in the DFW metropolitan and surrounding areas 
to drive these framework and find the influence of white spaces on network costs in these representative areas. 
%
%Further, we investigate different 
%band combinations in two population densities to show that greater access to white space 
%channels have greater total savings of mesh nodes when the total number of channels used 
%in the network is fixed (i.e., given a total number of allowable WiFi and white space channels). 
Through extensive analysis across varying population density and channel combinations across bands, 
we show that white space bands can reduce the number of access points by 1650\%
and 660\% in rural and sparse urban areas versus the same cost savings are not achieved in dense urban 
and downtown type area. As the population and spectrum utilization increase, the cost savings of 
white space bands diminish to the point that WiFi-only channel combinations can be optimal.
The simulation shows that our BPS algorithm can achieve 180\% the served user demand versus previous 
multi-channel, multi-radio solutions in multiband scenarios, since we leverage diverse propagation 
characteristics offered by WiFi and white space bands. Moreover,we quantify the degree to which the joint 
use of these bands can improve the served user demand. Our BPS algorithm shows that WhiteMesh topologies 
can achieve up to 160\% of the served traffic flow of similar WiFi or white-space-only configurations.
% Future work
%In the future, we will consider the heterogeneous access points and traffic demand scenarios
%in wireless network deployments.


